---
title: "Caitlin Clark Career Stats Insights"
author: "Mia Fitzgerald & Angela McMahon"
date: "4/8/2026"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=TRUE, message=TRUE)
library(tidyverse)
library(knitr)

```


## Introduction 


Describe your dataset and its topic, and summarize what it measures,
and its sample size. Include a link to your data set.


### Data Import


Discuss any steps you took to load and clean the data in as non-technical
terms as you can. Include your data import/cleaning chunk here, but keep code folding closed by
default. If you needed to process the date data, keep that code here as well.

```{r}
library(readr)
library(tidyr)
CC_career_stats <- read_csv("CC_career_stats.csv")
head(CC_career_stats)
CC_career_stats%>%
  mutate(Away=if_else(Away=="@","Yes","No"), Away=coalesce(Away, "No"))%>%
  mutate(GS=if_else(GS=="*","Yes", "No"))%>%   #did not fill NA because all are Yes 
  separate_wider_delim(
    cols=Result,
    delim=" ",
    names=c("Result", "Points"),
    too_few="align_start",
    too_many="merge"
  )

```


### Exploratory Data Analysis 
### add subsections with informative titles


a. Explore your variables. You don’t need to include every plot/table you try out, but include the
clearest, most illustrating ones. At the very least, explore your date variable, quantitative variable,
and the bivariate relationship between them.

b. Depending on the scale of your date data, explore any notable patterns (number of observations,
value of quantitative variable, etc.) that show up relating to day of the week, month of the year,
etc.

c. Similarly, depending on the scale of your date data, explore any notable patterns that show up
relating to weekend versus weekday, season/quarter of the year, etc. 


### Conclusion 


Summarize your findings.



