---
title: "Caitlin Clark Career Stats Insights"
author: "Mia Fitzgerald & Angela McMahon"
date: "4/8/2026"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=TRUE, message=TRUE)
library(tidyverse)
library(lubridate)
library(knitr)
library(ggplot2)

```


## Introduction 


Describe your dataset and its topic, and summarize what it measures,
and its sample size. Include a link to your data set.


### Data Import

From the data set table, made the data into a csv file format. Copied and pasted this into a text file. Here the column names including symbols like % and a missing column name were altered to correspond well with R. This file was saved as a csv file and imported into the R markdown file. Once the data set was loaded in, for both the columns 'Away' and 'Games Started' the 'Yes' notation was symbolized by a character. So, this was changed to a Yes or No input. The 'Results' column also had both W or L for wins and losses and the score at the end of each game. This column was separated into 'Results' which records wins 'W' and losses 'L', then 'Points' which is the score at the end of the game for each team. To fill missing values and merge we used too_few/too_many inside of widening the column. Once these adjustments were made, this new, tidy data set was restored. 


```{r}
library(readr)
library(tidyr)
CC_career_stats <- read_csv("CC_career_stats.csv")
CC_career_stats <- CC_career_stats%>%
  mutate(Away=if_else(Away=="@","Yes","No"), Away=coalesce(Away, "No"))%>%
  mutate(GS=if_else(GS=="*","Yes", "No"))%>%
  separate_wider_delim(
    cols=Result,
    delim=" ",
    names=c("Result", "Points"),
    too_few="align_start",
    too_many="merge"
  )
head(CC_career_stats)

```


### Exploratory Data Analysis 
### add subsections with informative titles


a. Explore your variables. You don’t need to include every plot/table you try out, but include the
clearest, most illustrating ones. At the very least, explore your date variable, quantitative variable,
and the bivariate relationship between them.


```{r}
#instances where she shot over 50 percent from the 3 over her career 
CC_career_stats %>%
  filter(!is.na(Date)) %>%
  ggplot(aes(x = yday(Date), y = `3PP`)) +
  geom_point(aes(color = `3PP` > 0.5), alpha = 0.6) +
  facet_wrap(~year(Date))+
  labs(x="Days of a year", y="Three-point percentage")
  
#instances where she shot over 50 percent from the FT over her career (A LOT)
CC_career_stats %>%
  filter(!is.na(Date), !is.na(FTP)) %>%
  ggplot(aes(x = yday(Date), y = FTP)) +
  geom_point(alpha = 0.5) 

# home or away vs AST (higher rate when AWAY)
CC_career_stats %>%
  filter(AST < 100) %>% #outlier removed
  ggplot(aes(x = Away, y = AST)) + 
  geom_boxplot(alpha = 0.7)+
  labs(y="Assists")

# Instances where CC scored over 35 points in her career 
CC_career_stats %>%
  filter(PTS < 100) %>%
  ggplot(aes(x = Date, y = PTS, color = PTS > 35)) +
  geom_point(alpha = 0.6)+
  labs(y="Points")

#Difference in counts of rebound type throughout her career (MORE DRB)
CC_career_stats %>%
  pivot_longer(cols = c(ORB, DRB), names_to = "Rebound_Type", values_to = "Count") %>%
  filter(Count < 20) %>%
  ggplot(aes(x = yday(Date), y = Count)) +
  geom_point(alpha = 0.5) + 
  facet_wrap(~Rebound_Type)

CC_career_stats %>%
  pivot_longer(cols = c(ORB, DRB), names_to = "Rebound_Type", values_to = "Count") %>%
  filter(Count < 20)%>%
  ggplot(aes(x=Count))+
  geom_histogram(binwidth=1)+
  facet_wrap(~Rebound_Type)+
  labs(x="Number of Rebounds", y="Number of Occasions")

#Dates of limited turnovers (0 and 1) in her career from end of 2020-2021 to 2023-2024 seasons
CC_career_stats %>%
  select(Date, TOV, FGP, Result) %>%
  filter(TOV == 0 | TOV == 1) %>%
  kable()

#Games where she has more than 12 assists lead to wins
CC_career_stats %>%
  select(TOV, AST, Date, Result) %>%
  filter(!is.na(Date), AST > 12) %>%
  kable()

#After 11 AST there are no games lost
CC_career_stats %>%
  filter(AST < 100) %>%
  mutate(Season = ifelse(month(Date) >= 10, 
                         year(Date), 
                         year(Date) - 1)) %>%
  ggplot(aes(x = Date, y = AST, color = Result, group = interaction(Result, Season))) +
  geom_line() + 
  facet_wrap(~Season, scales = "free_x")

```



b. Depending on the scale of your date data, explore any notable patterns (number of observations,
value of quantitative variable, etc.) that show up relating to day of the week, month of the year,
etc.

c. Similarly, depending on the scale of your date data, explore any notable patterns that show up
relating to weekend versus weekday, season/quarter of the year, etc. 


### Conclusion 


Summarize your findings.



